### 2018-01-22 Chapters/Chapter01.tex
Update on Overleaf.

Bitcoin uses two hash functions: HASH-160 and HASH-256. The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}. The latter, has $n=256$ and applies \textit{SHA-256} twice.

@@

\section{Bitcoin mining}

Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function. The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given number. The lesser the $A$, the harder to find a new block. In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.

@@

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\tau H} \right)^{tH}$. But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.

@@

	As $\log(1 + \frac{1}{M}) > 0$, $1 < 1 + \log(1 + \frac{1}{M})$.

@@

For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ --- or $[0.51, 30.76]$ in minutes. Thus, 90\% of the time between blocks blocks are between 30 seconds and 31 minutes.

@@

\begin{figure}[ht]
\centering\includegraphics[height=5cm]{./images01/time-6-blocks.pdf}
\caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
\end{figure}

@@

\begin{cor}
	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximatelly multiplies their probability of finding the next block by $m$.
\end{cor}
\begin{proof}
	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
\end{proof}

@@

%Then, $\mathbf{P}(t_0 < T < t_1) = \mathbf{P}(t_0 < X/H < t_1) = \mathbf{P}(t_0 H < X < t_1 H) = [1 - (1-p)^{t_1 H}] - [1 - (1-p)^{t_0 H}] = (1-p)^{t_0 H} - (1-p)^{t_1 H}$.

@@

For example, if the network's total hash rate suddenly doubles ($\alpha = 2$), then $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved. On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled. It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, and then the adjustment will occur and everything will be back to the normal 10 minutes between blocks.

@@

\begin{theorem}
	$$a > 1, x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
\end{theorem}
\begin{proof}
$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$. As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$. Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$. Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
\end{proof}

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$. So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes. If the hash rate had suddenly doubled, the average time between blocks would be 300.

@@

\subsection{Hash rate changing linearly}

@@

@@

Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
Let $y = bx+c \Rightarrow dy = bdx$.

@@

As $\frac{d}{ds}c = \frac{10\sqrt{3}}{a}$,

\begin{align*}
\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{20\sqrt{3\pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{10\sqrt{3\pi}s}{2ab} \\
\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
\end{align*}

It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = \tau = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.

@@

Attackers are not part of the mining network, because they are trying to generate a different longest chain.

@@

As \cite{nakamoto2008bitcoin} have done, we will also model the attack using the Gambler's Ruin. In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$. The rounds are independent. The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money. Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:

@@

The gambler wins \$1 is the same as the network find a new block, the gambler losses \$1 is the same as the attacker find a new block. The initial \$$k$ is the same as the number of blocks the attacker is behind the network. Thus, the gambler losses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.

@@

But this scenario seems not to be the optimal attack, because the attacker has waited the $k$ confirmations before starting the attack. A better approach would be to start attacking just after propagating the transaction. In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations. So, we have to model the probabilities before the network has found the $k$ block. Then, if the attacker has more blocks than the network, he has successfully attacked. Otherwise, we got back to the previous model, in which the attacker must still find more blocks.

\begin{theorem}
	Let $a_{ij}$ be the probability of moving from state $(i, j)$ to state $(s, k)$ in the Markov chain.
	$$a_{ij} = \binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}.$$
\end{theorem}
\begin{proof}

In order to go from $(i, j)$ to $(s, k)$, the attacker must find exacly $i-s$ blocks, and the network must find exactly $k-j$ blocks. It is as they would be walking the grid from the point $(i, j)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}. There are many possible paths and all of them occurs with probability $(1-p)^{s-i} p^{k-j}$.

The walking ends when $(\cdot, k$) is reach, i.e., when the network finds $k$ blocks, no matter how many blocks the attacker has found. So, it is not allowed to walk through the line $(\cdot, k)$. Thus, the number of paths between $(i, j)$ and $(s, k)$ going only up or right, without going through the line $(\cdot, k)$ is exactly the number of paths between $(i, j)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s-i$ right movements ($r$) and $k-j-1$ up movements ($u$). The number of permutations is $\frac{(k-j-1+s-i)!}{(s-i)!(k-j-1)!} = \binom{(k-j-1)+(s-i)}{s-i}$ because there are $s-i$ repetitions of the element $r$ and $k-j-1$ repetitions of the element $u$.

Finally, the probability is $\binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}$.

@@

    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
  \end{tikzpicture}
  \caption{Both the attacker and the network are mining. Each step up is a new block found by the network with probability $p$. Each step right is a new block found by the attacker with probability $1-p$. It ends when the network finds $k$ blocks --- in this example, $k=6$. The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$. Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network. In the red path, the attacker still have to catch up 3 blocks to have a successfull attack, which happens with probability $\rho^3$, if $p < 0.5$.}
  \label{figure:bitcoin-attack-paths}
\end{figure}
\end{proof}

\begin{cor}
	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
\end{cor}
\begin{proof}
	This probability $\mathbf{P}(S = s) = a_{00}$.
\end{proof}

Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.

@@

Let's assume that all transactions have exactly the same size $S$ and pays exactly the same fee. If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transaction in each block. Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single zero of $z^s(1+m(1-z))-1$ with $|z_s|>1$. The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$. Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$. The probabilities form a simple geometric series, and so they are exponentially decreasing. We may interpret it as a stable system. In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\tau$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$. Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.

>>>>

Bitcoin uses two hash functions: HASH-160 and HASH-256. The first has $n=160$ and consists of the composition of \textit{SHA-256} and \textit{RIPEMD-160}. The latter has $n=256$ and applies \textit{SHA-256} twice.

@@

\section{Mining one block}

Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin \textit{HASH-256} function. The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given threshold. The smaller the $A$, the harder to find a new block. In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.

@@

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\tau H} \right)^{tH}$. But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\tau}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\tau}$.

@@

	As $\log(1 + \frac{1}{M}) > 0$ and $1 < 1 + \log(1 + \frac{1}{M})$.

@@

For instance, if $\alpha = 10\%$, then $t_0 = 30.77$ and $t_1 = 1797.44$ (or $[0.51, 30.76]$ in minutes). Thus, 90\% of the time the intervals between blocks are between 30 seconds and 30 minutes.

The fact that the time between blocks follows an exponential distribution with $\lambda = 1/\tau = pH$ may be used to estimate the total network's hash rate (or a miner's hash rate). For further information, see \cite{ozisikestimation}.

\section{Mining several blocks}

@@

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/time-6-blocks.pdf}
\caption{Probability density function of $Y_6$, i.e., probability of finding 6 blocks after time $t$. The shaded areas shows the lower 5\% and upper 5\% of the pdf.\label{fig-bitcoin-time-6-blocks}}
\end{figure}

\section{Mining for a miner}

@@

\begin{cor}
	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximately multiplies their probability of finding the next block by $m$.
\end{cor}
\begin{proof}
	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
\end{proof}

@@

\section{Orphan blocks}

@@

For example, if the network's total hash rate suddenly doubles ($\alpha = 2$), then $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved. On the other side, if the network's total hash rate suddenly halves ($\alpha = 0.5$), then $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled. It is an important conclusion, since it shows that even if half of the network stops mining, it will only double the time to the next difficulty adjustment, i.e., the time between blocks will be 20 minutes for, at most, the next 29 days, at which point the adjustment will occur and everything will be back to the normal 10 minutes between blocks.

@@

\begin{theorem}
	$$a > 1 \text{ and } x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
\end{theorem}
\begin{proof}
$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$. As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$. Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$. Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
\end{proof}

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, then the parameters would be $\lambda = 1/600$, $a=2$, $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$, $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01 \Leftarrow b > 0.000163690 \Leftarrow n < 10.1818$. So, for $n=10$, then $b=0.000166666$ and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. In other words, during the next week, the average time between blocks will be 9 minutes and 17 seconds, instead of the normal 10 minutes. If the hash rate had suddenly doubled, the average time between blocks would be 5 minutes.

@@

\subsection{Piecewise linear model of hash rate change}

@@

It is well known that $\mathbf{E}(T) = \int_{0}^{\infty} 1 - \mathbf{P}(T \leq t) dt$. Thus, replacing $y = \frac{a^2 t + 1}{a \sqrt{2 \tau}}$, and using the fact that $\int_0^\infty e^{-x^2} dx = \sqrt{pi}/2 \erf(x)$, we have:

\begin{align}
\mathbf{E}(T)|_{t_1}^{t_2} &= \int_{t_1}^{t_2} \exp \left( - \frac{t + a^2 t^2/2}{\tau} \right) dt \nonumber \\
	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \int_{y_1}^{y_2}  \exp(-y^2) dy \nonumber \\
	&= \frac{\sqrt{2 \tau}}{a} \exp \left( \frac{1}{2a^2 \tau} \right) \frac{\sqrt{\pi}}{2} [\erf(y_1) - \erf(y_2)] \nonumber \\
	&= \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[\erf(y_2) - \erf(y_1) \right]
\end{align}

Where $y_1 = \frac{a^2 t_1 + 1}{a \sqrt{2 \tau}}$ and $y_2 = \frac{a^2 t_2 + 1}{a \sqrt{2 \tau}}$.

Thus, $\mathbf{E}(T) = \mathbf{E}(T)|_0^\infty$. When $t_1 = 0 \Rightarrow y_1 = \frac{1}{2 \sqrt{2 \tau}}$ and $t_2 \rightarrow \infty \Rightarrow y_2 \rightarrow \infty \Rightarrow \erf(y_2) = 1$, then:

$$
\mathbf{E}(T) = \frac{\sqrt{2 \pi \tau}}{2a} \exp \left( \frac{1}{2a^2 \tau} \right) \left[1 - \erf\left(\frac{1}{a \sqrt{2 \tau}}\right)\right]
$$

The function $\mathbf{E}(T)|_{t_1}^{t_2}$ may be used to a piecewise linear analysis of any hash rate change. Let's analyze the hash doubling in one week. Then, $u(1 \text{week}) = u(604,800) = 2$, thus $a^2 = \frac{1}{604,800} \Rightarrow a = \frac{1}{120 \sqrt{42}}$. Let's partition the interval $[0, 1 \text{week}]$ in hourly intervals, i.e., ($t_0$, $t_1$, $t_2$, \dots, $t_168$), where $t_i = i \cdot 1 \text{hour} = i \cdot 3600$. For each interval $[t_i, t_{i+1}]$, we evaluate $\mathbf{E}(T)|_{t_i}^{t_{i+1}}$. The result is presented in Figure .

---

@@

Let $b=\frac{a}{\sqrt(2 \tau)}$ and $c = \frac{\sqrt(2 \tau)}{2a} \left( s + \frac{1}{\tau} \right)$, then $sx + \frac{x+a^2x^2/2}{\tau} = (bx + c)^2 - c^2$.
Let $y = bx+c \Rightarrow dy = bdx$.

@@

As $\frac{d}{ds}c = \frac{\sqrt{2 \tau}}{2a}$,

\begin{align*}
\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{\sqrt{2 \tau \pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{\sqrt{2 \tau \pi}s}{4ab} \\
\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
\mathbf{E}[T] &= \frac{\sqrt{2 \tau \pi}}{2a} e^{\frac{1}{2 \tau a^2}} \erfc\left(\frac{1}{a \sqrt{2 \tau}}\right) \\
\end{align*}

@@

There are many possible ways to attack the Bitcoin (REF). In this section, we are interested in a particular attack: the submarine attack.

In the submarine attack, the attacker's goal is to spend some Bitcoins and, after a while, get their Bitcoins back. Thus, the attacker sends some funds to the victim, they wait 6 confirmations, and, finally, the victim confirms the funds. But, then, the attacker get their funds back propagating new blocks which accept a double spending transaction. In order to work, the attacker must propagate more blocks than the network has mined in the same period, becoming the longest chain. Hence, we would like to understand what the odds are that the attacker will be successful.

In order to maximize their odds, the attacker must start to mine the new blocks as soon as they send the funds to the victim. In this moment, it starts to mine in the head of the blockchain, just like the rest of the network. So, the attacker and the network are in exactly the same point.

@@

As \cite{nakamoto2008bitcoin} has done, we will also model the attack using the Gambler's Ruin. In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$. The rounds are independent. The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money. Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:

@@

The gambler winning \$1 is the same as the network finding a new block, the gambler losing \$1 is the same as the attacker finding a new block. The initial \$$k$ is the same as the number of blocks the attacker is behind the network. Thus, the gambler loses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.

@@

But this scenario seems not to be the optimal attack, because the attacker has waited $k$ confirmations before starting the attack. A better approach would be to start attacking just after propagating the transaction. In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations. So, we have to model the probabilities before the network has found the $k$ block. Then, if the attacker has more blocks than the network, he has successfully attacked. Otherwise, we return to the previous model, in which the attacker must still find more blocks.

\begin{theorem}
	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
\end{theorem}
\begin{proof}
The attacker must find exactly $s$ blocks while the network must find exactly $k$ blocks. It is as they would be walking the grid from the point $(0, 0)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}. When the attacker finds a block, it would be a movement to the right. When the network finds a block, it would be an upward movement. No matter the order which the blocks are found, all the paths occur with probability $(1-p)^s p^k$.

The walking ends when $(\cdot, k$) is reached, i.e., when the network finds $k$ blocks, regardless of how many blocks the attacker has found -- i.e., it is not allowed to walk above the line $(\cdot, k)$. Thus, the number of paths between $(0, 0)$ and $(s, k)$ moving only upward or to the right, without going into the line $(\cdot, k)$ is exactly the number of paths between $(0, 0)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s$ movements to the right ($r$) and $k-1$ upward movements ($u$). This number of permutations is $\frac{(k-1+s)!}{s!(k-1)!} = \binom{(k-1)+(s)}{s}$ because there are $s$ repetitions of the element $r$ and $k-1$ repetitions of the element $u$.

Finally, the probability is $\binom{k+s-1}{s} (1-p)^{s} p^{k}$.

@@

    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
  \end{tikzpicture}
  \caption{Both the attacker and the network are mining. Each step up is a new block found by the network with probability $p$. Each step right is a new block found by the attacker with probability $1-p$. It ends when the network finds $k$ blocks --- in this example, $k=6$. The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$. Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network. In the red path, the attacker still have to catch up 3 blocks to have a successful attack, which happens with probability $\rho^3$, if $p < 0.5$.}
  \label{figure:bitcoin-attack-paths}
\end{figure}
\end{proof}

Assuming that the attacker starts mining just after publishing the victim's transaction, the probability of the attacker will have found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.

@@

Let's assume that all transactions have exactly the same size $S$ and pay exactly the same fee. If the Bitcoin block's maximum size is $M$, there would be room for $s = \lfloor M/S \rfloor$ transactions in each block. Using the results from \citet{bailey1954queueing}, we have found that $\pi_n = \frac{z_s-1}{z_s^{n+1}}$ is the probability of having $n$ unconfirmed transactions in the pool subjected to $s > m$, where $m=\frac{\lambda_{\text{TX}}}{\lambda_{\text{blocks}}}$ and $z_s$ is the single root of the polynomial $z^s(1+m(1-z))-1$ with $|z_s|>1$. The average size of the unconfirmed transaction pool is $\mathbf{E}(\pi) = \frac{1}{z_s - 1}$. Thus, $s > m \Rightarrow \lambda_{\text{TX}} < s \lambda_{\text{blocks}}$. The probabilities form a simple geometric series, and so they are exponentially decreasing. We may interpret it as a stable system, i.e., the unconfirmed transactions pool size is finite. In the Bitcoin's network, $\lambda_{\text{blocks}} = 1/\tau$ and the average number of transactions per block is $s=2,250$, so, this solution is valid when $\lambda_{\text{TX}} < 2,250/600 = 3.75$. Hence, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$, the unconfirmed transaction pool starts to grow indefinitely.
